```{r}
require(purrr)
require(hdf5r)
require(inspre)
require(ggplot2)
require(tidyr)
require(tibble)
require(dplyr)
require(readr)
require(igraph)
require(ggupset)
require(betareg)
```


```{r}
k562_norm_sc_fn = '/gpfs/commons/groups/knowles_lab/gwps/data/K562_essential_normalized_singlecell_01.h5ad'
save_dag_res = '/gpfs/commons/groups/knowles_lab/gwps/saved_rdata/gwps_res_dag.Rdata'
save_guide_res = '/gpfs/commons/groups/knowles_lab/gwps/saved_rdata/gwps_guide_data.Rdata'
save_gene_data = '/gpfs/commons/groups/knowles_lab/gwps/saved_rdata/gene_data.Rdata'
save_graph_data = '/gpfs/commons/groups/knowles_lab/gwps/saved_rdata/gwps_graph_data.Rdata'
```


```{r}
min_nz <- 0.03/2
load(save_dag_res)
# best_lambda <- which.min(res_dag$eps_hat_G)
best_lambda <- 6 # Big drop in eps_hat_R, good rho_G, largest D_hat > 0.01
R_hat <- res_dag$R_hat
SE_hat <- res_dag$SE_hat
G_hat <- res_dag$G_hat[,,best_lambda]
G_hat[abs(c(G_hat)) < min_nz] <- 0
U_hat <- res_dag$U[,,best_lambda]

hfile_sc <- hdf5r::H5File$new(k562_norm_sc_fn, "r")
var <- inspre::parse_hdf5_df(hfile_sc, 'var')

load(save_guide_res)
```


```{r}
genes <- colnames(G_hat)
rownames(R_hat) <- genes
rownames(G_hat) <- genes
rownames(U_hat) <- genes
colnames(U_hat) <- genes

R_long <- tidyr::pivot_longer(tibble::as_tibble(R_hat, rownames="Exposure"), -Exposure,
                              names_to="Outcome", values_to = "R_hat")
SE_long <- tidyr::pivot_longer(tibble::as_tibble(SE_hat, rownames="Exposure"), -Exposure,
                              names_to="Outcome", values_to = "SE_hat")
G_long <- tidyr::pivot_longer(tibble::as_tibble(G_hat, rownames="Exposure"), -Exposure,
                              names_to="Outcome", values_to = "G_hat")
U_long <- tidyr::pivot_longer(tibble::as_tibble(U_hat, rownames = "Exposure"), -Exposure,
                              names_to = "Outcome", values_to = "U_hat")
graph_data <- R_long
graph_data$SE_hat <- SE_long$SE_hat
graph_data$G_hat <- G_long$G_hat
graph_data$U_hat <- U_long$U_hat

graph_data <- graph_data %>%
  dplyr::mutate(Z = R_hat/SE_hat, p = 2*(1-pnorm(abs(Z))), p_fdr = p.adjust(p, method='fdr'))
```


```{r}
print(summary(abs(graph_data$R_hat)))
print(summary(abs(graph_data$G_hat)))
print(summary(abs(graph_data$G_hat[abs(graph_data$G_hat) > 0.0])))
print(summary(abs(graph_data$U_hat)))
graph_data %>% ggplot(aes(x=abs(G_hat))) + geom_histogram()
graph_data %>% filter(abs(G_hat) > 0.0) %>% ggplot(aes(x=abs(G_hat))) + geom_histogram()
graph_data %>% ggplot(aes(x=abs(R_hat))) + geom_histogram()
graph_data %>% ggplot(aes(x=abs(U_hat))) + geom_histogram()
```



```{r, fig.widith=3.25, fig.height=3.25}
plot_order <- hclust(dist(U_hat))$order

graph_data <- dplyr::mutate(graph_data, R_plot=if_else(Exposure==Outcome, 0, R_hat), U_plot=if_else(Exposure==Outcome, 0, U_hat)) 
```


```{r, fig.width=2.72, fig.asp=1.08}
ggplot(graph_data, aes(y=Exposure, x=Outcome)) + # geom_tile(aes(fill=R_plot)) +
  geom_raster(aes(fill=sign(R_plot)*sqrt(abs(R_plot)))) +
  scale_x_discrete(limits = genes[(plot_order)], labels = NULL) +
  scale_y_discrete(limits = genes[rev(plot_order)], labels = NULL) +
  scale_fill_gradient2(limits = c(-1, 1))  +
  labs(x = "Outcome", y = "Exposure", title = "Observed marginal ACE (R)") +
  theme(plot.title = element_text(size=10), axis.ticks=element_blank(),
        axis.title = element_text(size=10), legend.title = element_text(size=10),
        plot.margin = unit(c(0.5,0.5,0.5,0.5), 'mm')) +
  guides(fill=FALSE) + coord_fixed()
```


```{r, fig.width=2.53, fig.asp=1.16}
ggplot(graph_data, aes(y=Exposure, x=Outcome)) + # geom_tile(aes(fill=U_plot)) +
  geom_raster(aes(fill=sign(U_hat)*sqrt(abs(U_hat)))) +
  scale_x_discrete(limits = genes[(plot_order)], labels = NULL) +
  scale_y_discrete(limits = genes[rev(plot_order)], labels = NULL) +
  scale_fill_gradient2(limits = c(-1, 1))  +
  labs(x = "Outcome", y = NULL, title = "Shrunk marginal ACE (U)") +
  theme(plot.title = element_text(size=10), axis.ticks=element_blank(),
        axis.title = element_text(size=10), legend.title = element_text(size=10),
        plot.margin = unit(c(0.5,0.5,0.5,0.5), 'mm')) +
  guides(fill=FALSE) + coord_fixed()
```


```{r, fig.width=3.26, fig.asp=0.9}
ggplot(graph_data, aes(y=Exposure, x=Outcome)) + # geom_tile(aes(fill=G_hat)) + 
  geom_tile(aes(fill=sign(G_hat)*sqrt(abs(G_hat)))) +
  scale_x_discrete(limits = genes[(plot_order)], labels = NULL) +
  scale_y_discrete(limits = genes[rev(plot_order)], labels = NULL) +
  scale_fill_gradient2(limits = c(-1, 1)) + #, breaks = c(-1, 0, 1)) +
  labs(x = "Outcome", y = NULL, fill = "Effect", title = "Inferred direct causal effect (G)") +
  theme(plot.title = element_text(size=10), axis.ticks=element_blank(),
        axis.title = element_text(size=10), legend.title = element_text(size=9),
        legend.key.size = unit(0.4, 'cm'), legend.text = element_text(size=9),
        plot.margin = unit(c(0.5,0,0.5,0.5), 'mm')) +
  coord_fixed()
```



```{r}
gene_data <- as_tibble_col(genes, column_name = "gene")
gene_data <- gene_data %>% dplyr::left_join(
  select(var, gene_id, gene_name, cv, mean, std), by=c("gene" = "gene_id")) %>%
  dplyr::left_join(select(guide_effects, target, inst_beta, inst_cor, n), by=c("gene"="target"))
gene_data <- gene_data %>% dplyr::left_join(enframe(genes[plot_order], name="order", value="gene"), by="gene")
```


```{r}
scale_factor <- 2
graph <- make_igraph(G_hat, min_edge_value = 0.05, scale_factor = scale_factor)

distances <- igraph::distances(graph, mode = "out")
distances_long <- tidyr::pivot_longer(
  as_tibble(distances, rownames = "Exposure"), -Exposure,
  names_to = "Outcome", values_to = "log_scale_dist")

graph_data$G_path_scaled <- 1/exp(distances_long$log_scale_dist)
graph_data$path_length <- purrr::pmap_int(
  select(graph_data, Exposure, Outcome, G_path_scaled),
  function(Exposure, Outcome, G_path_scaled){
    # split <- strsplit(Exposure, "_", fixed=TRUE)
    # as.integer(ifelse(G_path > 0, length(igraph::shortest_paths(graph, split[[1]][length(split[[1]])], Outcome, mode = "out")$vpath[[1]]) - 1, NA))
    as.integer(ifelse(G_path_scaled > 0,
                      length(igraph::shortest_paths(graph, Exposure, Outcome, mode = "out")$vpath[[1]]) - 1, NA))
  }
)

graph_data <- graph_data %>% dplyr::mutate(G_path = ifelse(G_path_scaled > 0, G_path_scaled*(scale_factor**path_length), 0),
                       eff_explained = ifelse((Exposure != Outcome) & (G_path > 0), abs(G_path/U_norm), NA))
gene_data$out_deg <- igraph::degree(graph, mode = "out")[gene_data$gene]
gene_data$in_deg <- igraph::degree(graph, mode = "in")[gene_data$gene]
gene_data$deg <- igraph::degree(graph, mode = "all")[gene_data$gene]
gene_data$bt_centrality <- igraph::betweenness(graph, normalized = TRUE)
gene_data$eg_centrality <- igraph::eigen_centrality(graph)$vector
```


```{r}
save(gene_data, file=save_gene_data)
save(graph_data, file=save_graph_data)
```


```{r}
load(save_gene_data)
load(save_graph_data)
```


```{r, fig.height=2, fig.width=2.8}

graph_data %>% filter(!is.na(path_length), Exposure != Outcome) %>%
  ggplot(aes(x=path_length)) + geom_bar(aes(y=..count../sum(..count..))) +
  labs( title = "Shortest path between nodes (all)", x = "Path length", y = "Proportion") + 
  theme_classic() +
  theme(plot.title = element_text(size=10),  axis.title = element_text(size=10)) +
  scale_x_continuous(breaks = scales::pretty_breaks(n=6))
 
graph_data %>% filter(!is.na(path_length), Exposure != Outcome, p_fdr < 0.05) %>%
  ggplot(aes(x=path_length)) + geom_bar(aes(y=..count../sum(..count..))) +
  labs( title = "Shortest path between nodes", x = "Path length", y = "Proportion") + 
  theme_classic() +
  theme(plot.title = element_text(size=10), axis.title = element_text(size=10)) +
#        axis.ticks.y = element_blank(), axis.text.y=element_blank()) +
  scale_x_continuous(breaks = scales::pretty_breaks(n=6))


gene_data %>% ggplot(aes(x = in_deg)) + geom_histogram() + labs( x = "In-degree")

gene_data %>% ggplot(aes(x = out_deg)) + geom_histogram() + labs( x = "Out-degree")

print(summary(filter(graph_data, !is.na(eff_explained), p_fdr < 0.05)$eff_explained))
```
```{r, fig.height=2, fig.width=2.5}
graph_data %>% filter(!is.na(eff_explained), p_fdr < 0.05) %>% ggplot(aes(x=100*eff_explained)) +
  geom_histogram(aes(y=..count../sum(..count..)), bins = 50) + xlim(0, 150) +
  labs(title = "ACE explained by shortest path",
       x = "Percentage of effect explained", y = NULL) +
  theme_classic() + 
  theme(plot.title = element_text(size=10), axis.title = element_text(size=10)) + ylim(0, 0.3)

```


```{r}
print((sum(!is.na(graph_data$path_length))-788)/(nrow(graph_data)-788))
print(mean(graph_data$path_length, na.rm=T))
print(sd(graph_data$path_length, na.rm=T))
print(sd(graph_data$path_length, na.rm=T)/sqrt(length(graph_data$path_length)))
print(mean(filter(graph_data, p_fdr<0.05)$path_length, na.rm=T))
print(sd(filter(graph_data, p_fdr<0.05)$path_length, na.rm=T))
print(sd(filter(graph_data, p_fdr<0.05)$path_length, na.rm=T)/sqrt(nrow(filter(graph_data, p_fdr<0.05))))
print(median(filter(graph_data, p_fdr<0.05)$eff_explained, na.rm=T))
print(sum(filter(graph_data, p_fdr<0.05)$eff_explained > 1, na.rm=T))
```


```{r}
gene_data %>% ggplot(aes(out_deg, inst_beta)) + geom_point()
gene_data %>% ggplot(aes(out_deg, inst_cor)) + geom_point()
gene_data %>% ggplot(aes(in_deg, inst_cor)) + geom_point()
gene_data %>% ggplot(aes(cv, inst_beta)) + geom_point()
gene_data %>% ggplot(aes(mean, inst_beta)) + geom_point()
gene_data %>% ggplot(aes(mean, in_deg)) + geom_point()
gene_data %>% ggplot(aes(exp_norm, deg)) + geom_point()
gene_data %>% ggplot(aes(cv, out_deg)) + geom_point()
gene_data %>% ggplot(aes(mean, eg_centrality)) + geom_point()
gene_data %>% ggplot(aes(inst_cor, eg_centrality)) + geom_point()
gene_data %>% ggplot(aes(bt_centrality, eg_centrality)) + geom_point()
```

```{r, fig.height=2.13, fig.width=2.8}
gene_data %>% dplyr::arrange(desc(eg_centrality)) %>% head(10) %>%
  ggplot(aes(x=reorder(gene_name, eg_centrality, decreasing = TRUE), y=eg_centrality)) + geom_col() +
  labs(x = NULL, y = "Eigencentrality", title="Highest eigencentrality genes") + 
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1),
        plot.title = element_text(size=10),
        axis.title = element_text(size=10))

# gene_data %>% dplyr::arrange(desc(bt_centrality)) %>% head() %>% ggplot(aes(x=gene_name, y=bt_centrality)) + geom_col() + theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
# gene_data %>% dplyr::arrange(desc(out_deg)) %>% head() %>% ggplot(aes(x=gene_name, y=out_deg)) + geom_col() + theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
# gene_data %>% dplyr::arrange(desc(deg)) %>% head() %>% ggplot(aes(x=gene_name, y=deg)) + geom_col() + theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
```

```{r}
gene_annotations_MM <- readr::read_delim('/gpfs/commons/groups/knowles_lab/gwps/data/GeneAnnotation_MM.tsv')
gene_annotations_JM <- readr::read_delim('/gpfs/commons/groups/knowles_lab/gwps/data/MorrisScience2023_SupplementaryTable_refGeneAnnotations.txt')

gene_data <- gene_data %>% dplyr::left_join(gene_annotations_MM, by=c("gene_name" ="gene"))
gene_data <- gene_data %>% dplyr::left_join(gene_annotations_JM, by=c("gene_name" ="gene"))
gene_data <- mutate(gene_data, eg_cent_mod = if_else(
  eg_centrality==1, 0.999, if_else(eg_centrality==0, 0.001, eg_centrality)))
gene_data <- mutate(gene_data, bt_cent_mod = if_else(
  bt_centrality==1, 0.999, if_else(bt_centrality==0, 0.001, bt_centrality)))
```


# ```{r}
# calc_annot_stats <- function(column){
#   betareg_res <- summary(betareg::betareg(as.formula(paste("eg_cent_mod ~ mean + ", column)), data=gene_data))     
#   return(as.data.frame(as.list(betareg_res$coefficients$mean[3,]), row.names = column))
# }
# 
# MM_annot_res <- purrr::map(colnames(gene_annotations_MM)[2:19], calc_annot_stats) %>% list_rbind()
# MM_annot_res$p_adj <- p.adjust(MM_annot_res$Pr...z.., method='holm')
# as_tibble(MM_annot_res_nomean, rownames = "Annotation") %>% readr::write_csv(file = "/gpfs/commons/groups/knowles_lab/gwps/saved_rdata/mm_annot_res.csv")
# MM_annot_res
# ```


```{r}
calc_beta_res <- function(column){
  betareg_res <- betareg::betareg(as.formula(paste("eg_cent_mod ~ ", column)), data=gene_data)
  return(betareg_res)
}

get_coeff <- function(betareg_res){
  summary_res <- summary(betareg_res)
  column <- rownames(summary_res$coefficients$mean)[2]
  return(as.data.frame(as.list(summary_res$coefficients$mean[2,]), row.names = column))
}

# MM_annot_res_nomean <- purrr::map(colnames(gene_annotations_MM)[2:17], calc_annot_stats) %>% list_rbind()
# MM_annot_res_nomean$p_adj <- p.adjust(MM_annot_res_nomean$Pr...z.., method='holm')
# readr::write_csv(MM_annot_res_nomean, file = "/gpfs/commons/groups/knowles_lab/gwps/saved_rdata/mm_annot_res_nomean.csv")
#  MM_annot_res_nomean
all_beta_res <- purrr::map(colnames(gene_annotations_MM)[2:17], calc_beta_res)
MM_annot_res <- purrr::map(all_beta_res, get_coeff) %>% list_rbind()
names(all_beta_res) <- colnames(gene_annotations_MM)[2:17]
MM_annot_res$p_adj <- p.adjust(MM_annot_res$Pr...z.., method='holm')
readr::write_csv(MM_annot_res, file = "/gpfs/commons/groups/knowles_lab/gwps/saved_rdata/mm_annot_res.csv")
MM_annot_res %>% arrange(p_adj)
```


```{r, fig.width=2.7, fig.height=2}
# Plotting significant relationships
line_data = data.frame(x=seq(0, 1, .01), y=predict(all_beta_res$pTriplo, newdata = data.frame(pTriplo=seq(0, 1, .01))))
gene_data %>% ggplot(aes(x=pTriplo, y=eg_centrality)) + geom_point() +
  geom_line(data=line_data, aes(x=x, y=y), color='blue', linewidth=1) + theme_classic()

line_data = data.frame(x=seq(0.24, 0.84, .01), y=predict(all_beta_res$EDS, newdata = data.frame(EDS=seq(0.24, 0.84, .01))))
gene_data %>% ggplot(aes(x=EDS, y=eg_centrality)) + geom_point() +
  geom_line(data=line_data, aes(x=x, y=y), color='blue', linewidth=1) + theme_classic() + 
  labs(x = "Enhancer domain score", y = "Eigencentrality")

line_data = data.frame(x=seq(0, 7500, 1), y=predict(all_beta_res$n_ppis, newdata = data.frame(n_ppis=seq(0, 7500, 1))))
gene_data %>% ggplot(aes(x=n_ppis, y=eg_centrality)) + geom_point() +
  geom_line(data=line_data, aes(x=x, y=y), color='blue', linewidth=1)  + theme_classic()

line_data = data.frame(x=seq(0, 2, .01), y=predict(all_beta_res$gnomAD_LOEUF, newdata = data.frame(gnomAD_LOEUF=seq(0, 2, .01))))
gene_data %>% ggplot(aes(x=gnomAD_LOEUF, y=eg_centrality)) + geom_point() +
  geom_line(data=line_data, aes(x=x, y=y), color='blue', linewidth=1) + theme_classic()

line_data = data.frame(x=seq(-2.6, 2.3, .1), y=predict(all_beta_res$ExAC_DEL_Z, newdata = data.frame(ExAC_DEL_Z=seq(-2.6, 2.3, .1))))
gene_data %>% ggplot(aes(x=ExAC_DEL_Z, y=eg_centrality)) + geom_point() +
  geom_line(data=line_data, aes(x=x, y=y), color='blue', linewidth=1) + theme_classic()

line_data = data.frame(x=seq(0, 2, .01), y=predict(all_beta_res$gnomAD_MisOEUF, newdata = data.frame(gnomAD_MisOEUF=seq(0, 2, .01))))
gene_data %>% ggplot(aes(x=gnomAD_MisOEUF, y=eg_centrality)) + geom_point() +
  geom_line(data=line_data, aes(x=x, y=y), color='blue', linewidth=1) + theme_classic()

line_data = data.frame(x=seq(0, 1, .01), y=predict(all_beta_res$HI_Index, newdata = data.frame(HI_Index=seq(0, 1, .01))))
gene_data %>% ggplot(aes(x=HI_Index, y=eg_centrality)) + geom_point() +
  geom_line(data=line_data, aes(x=x, y=y), color='blue', linewidth=1) + theme_classic()

line_data = data.frame(x=seq(-2.9, 2.2, .1), y=predict(all_beta_res$CCDG_DEL_Score, newdata = data.frame(CCDG_DEL_Score=seq(-2.9, 2.2, .1))))
gene_data %>% ggplot(aes(x=CCDG_DEL_Score, y=eg_centrality)) + geom_point() +
  geom_line(data=line_data, aes(x=x, y=y), color='blue', linewidth=1) + theme_classic()
```


```{r}
plt_genes <- filter(gene_data, Platelets == TRUE, (out_deg > 0 | in_deg > 0))$gene
ret_genes <- filter(gene_data, Reticulocytes == TRUE, (out_deg > 0 | in_deg > 0))$gene
wbc_genes <- filter(gene_data, WBCs == TRUE, (out_deg > 0 | in_deg > 0))$gene
rbc_genes <- filter(gene_data, RBCs == TRUE, (out_deg > 0 | in_deg > 0))$gene
all_blood_genes <- union(union(plt_genes, ret_genes), union(rbc_genes, wbc_genes))
```


```{r, fig.width=5in, fig.asp=3}
make_upset_res <- function(gene){
  res = c()
  if(gene %in% plt_genes) res = c(res, "PLT")
  if(gene %in% wbc_genes) res = c(res, "WBC")
  if(gene %in% rbc_genes) res = c(res, "RBC")
  if(gene %in% ret_genes) res = c(res, "RET")
  return(res)
}

blood_gene_data <- data.frame(gene=all_blood_genes)
blood_gene_data$blood_traits <- purrr::map(blood_gene_data$gene, make_upset_res)
blood_gene_data %>% ggplot(aes(x = blood_traits)) + geom_bar() + scale_x_upset() + theme_classic() + labs(x = NULL, y = "Gene count")
```

# 
# gene_data <- gene_data %>% dplyr::mutate(G_rnk_plt = min_rank(desc(G_plt)),
#                                          G_rnk_wbc = min_rank(desc(G_wbc)),
#                                          G_rnk_rbc = min_rank(desc(G_rbc)),
#                                          G_rnk_ret = min_rank(desc(G_ret)))
# 
# gene_data <- gene_data %>% dplyr::mutate(U_rnk_plt = min_rank(desc(U_plt)),
#                                          U_rnk_wbc = min_rank(desc(U_wbc)),
#                                          U_rnk_rbc = min_rank(desc(U_rbc)),
#                                          U_rnk_ret = min_rank(desc(U_ret)))
# 
# gene_data$top_trait_G <- dplyr::select(gene_data, starts_with("G_rnk")) %>%
#   rename_with(~gsub("G_", "", .x)) %>% purrr::pmap(make_upset_res)
# gene_data$n_top_traits_G <- purrr::map_int(gene_data$top_trait_G, length)
# 
# gene_data$top_trait_U <- dplyr::select(gene_data, starts_with("U_rnk")) %>%
#   rename_with(~gsub("U_", "", .x)) %>% purrr::pmap(make_upset_res)
# gene_data$n_top_traits_U <- purrr::map_int(gene_data$top_trait_U, length)
# 
# gene_data %>% filter(n_top_traits_G > 0) %>% ggplot(aes(x = top_trait_G)) + geom_bar() + scale_x_upset()
# gene_data %>% filter(n_top_traits_U > 0) %>% ggplot(aes(x = top_trait_U)) + geom_bar() + scale_x_upset()
# ```


# ```{r, fig.width=8}
# cutoff <- 0.1
# U_plt <- rowSums(abs(U_hat[,colnames(U_hat) %in% plt_genes]) > cutoff)
# G_plt <- rowSums(abs(G_hat[,colnames(G_hat) %in% plt_genes]) > cutoff)
# 
# U_wbc <- rowSums(abs(U_hat[,colnames(U_hat) %in% wbc_genes]) > cutoff)
# G_wbc <- rowSums(abs(G_hat[,colnames(G_hat) %in% wbc_genes]) > cutoff)
# 
# U_rbc <- rowSums(abs(U_hat[,colnames(U_hat) %in% rbc_genes]) > cutoff)
# G_rbc <- rowSums(abs(G_hat[,colnames(G_hat) %in% rbc_genes]) > cutoff)
# 
# U_ret <- rowSums(abs(U_hat[,colnames(U_hat) %in% ret_genes]) > cutoff)
# G_ret <- rowSums(abs(G_hat[,colnames(G_hat) %in% ret_genes]) > cutoff)
# 
# gene_data$U_plt <- U_plt
# gene_data$G_plt <- G_plt
# gene_data$U_wbc <- U_wbc
# gene_data$G_wbc <- G_wbc
# gene_data$U_rbc <- U_rbc
# gene_data$G_rbc <- G_rbc
# gene_data$U_ret <- U_ret
# gene_data$G_ret <- G_ret
# ```
# 
# 
# ```{r}
# gene_data %>% dplyr::arrange(desc(G_plt)) %>% head() %>% ggplot(aes(x=gene_name, y=G_plt)) + geom_col() + theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
# 
# gene_data %>% dplyr::arrange(desc(U_plt)) %>% head() %>% ggplot(aes(x=gene_name, y=U_plt)) + geom_col() + theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
# 
# gene_data %>% dplyr::arrange(desc(G_wbc)) %>% head() %>% ggplot(aes(x=gene_name, y=G_wbc)) + geom_col() + theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
# 
# gene_data %>% dplyr::arrange(desc(U_wbc)) %>% head() %>% ggplot(aes(x=gene_name, y=U_wbc)) + geom_col() + theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
# 
# 
# gene_data %>% dplyr::arrange(desc(G_rbc)) %>% head() %>% ggplot(aes(x=gene_name, y=G_rbc)) + geom_col() + theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
# 
# gene_data %>% dplyr::arrange(desc(U_rbc)) %>% head() %>% ggplot(aes(x=gene_name, y=U_rbc)) + geom_col() + theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
# 
# 
# gene_data %>% dplyr::arrange(desc(G_ret)) %>% head() %>% ggplot(aes(x=gene_name, y=G_ret)) + geom_col() + theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
# 
# gene_data %>% dplyr::arrange(desc(U_ret)) %>% head() %>% ggplot(aes(x=gene_name, y=U_ret)) + geom_col() + theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
# 
# ```
# 
# 
# ```{r}
# make_upset_res <- function(rnk_plt, rnk_wbc, rnk_rbc, rnk_ret, cutoff=10){
#   res = c()
#   if(rnk_plt <= cutoff) res = c(res, "Platelet")
#   if(rnk_wbc <= cutoff) res = c(res, "WBC")
#   if(rnk_rbc <= cutoff) res = c(res, "RBC")
#   if(rnk_ret <= cutoff) res = c(res, "Reticulocyte")
#   return(res)
# }
# 
# gene_data <- gene_data %>% dplyr::mutate(G_rnk_plt = min_rank(desc(G_plt)),
#                                          G_rnk_wbc = min_rank(desc(G_wbc)),
#                                          G_rnk_rbc = min_rank(desc(G_rbc)),
#                                          G_rnk_ret = min_rank(desc(G_ret)))
# 
# gene_data <- gene_data %>% dplyr::mutate(U_rnk_plt = min_rank(desc(U_plt)),
#                                          U_rnk_wbc = min_rank(desc(U_wbc)),
#                                          U_rnk_rbc = min_rank(desc(U_rbc)),
#                                          U_rnk_ret = min_rank(desc(U_ret)))
# 
# gene_data$top_trait_G <- dplyr::select(gene_data, starts_with("G_rnk")) %>%
#   rename_with(~gsub("G_", "", .x)) %>% purrr::pmap(make_upset_res)
# gene_data$n_top_traits_G <- purrr::map_int(gene_data$top_trait_G, length)
# 
# gene_data$top_trait_U <- dplyr::select(gene_data, starts_with("U_rnk")) %>%
#   rename_with(~gsub("U_", "", .x)) %>% purrr::pmap(make_upset_res)
# gene_data$n_top_traits_U <- purrr::map_int(gene_data$top_trait_U, length)
# 
# gene_data %>% filter(n_top_traits_G > 0) %>% ggplot(aes(x = top_trait_G)) + geom_bar() + scale_x_upset()
# gene_data %>% filter(n_top_traits_U > 0) %>% ggplot(aes(x = top_trait_U)) + geom_bar() + scale_x_upset()
# ```
# 
# ```{r}
# gene_data %>% filter(top_trait_G == c("Platelet")) %>% select(c('gene_name', starts_with('G_rnk')))
# gene_data %>% filter(top_trait_G == c("Reticulocyte")) %>% select(c('gene_name', starts_with('G_rnk')))
# gene_data %>% filter(top_trait_G == c("RBC")) %>% select(c('gene_name', starts_with('G_rnk')))
# gene_data %>% filter(top_trait_G == c("WBC")) %>% select(c('gene_name', starts_with('G_rnk')))
# ```
# 
# 
# ```{r}
# gene_data %>% filter(top_trait_U == c("Platelet")) %>% select(c('gene_name', starts_with('U_rnk')))
# gene_data %>% filter(top_trait_U == c("Reticulocyte")) %>% select(c('gene_name', starts_with('U_rnk')))
# gene_data %>% filter(top_trait_U == c("RBC")) %>% select(c('gene_name', starts_with('U_rnk')))
# gene_data %>% filter(top_trait_U == c("WBC")) %>% select(c('gene_name', starts_with('U_rnk')))
# ```


```{r}
all_blood_genes <- union(union(plt_genes, ret_genes), union(rbc_genes, wbc_genes))

test_mean_effect <- function(gene){
  t_plt <- t.test(abs(filter(graph_data, Exposure == gene, Outcome %in% plt_genes)$U_hat),
                  abs(filter(graph_data, Exposure == gene, Outcome %in% 
                           setdiff(unique(c(wbc_genes, rbc_genes, ret_genes)), plt_genes))$U_hat), 
                  alternative = "greater")$statistic[['t']]
  t_ret <- t.test(abs(filter(graph_data, Exposure == gene, Outcome %in% ret_genes)$U_hat),
                  abs(filter(graph_data, Exposure == gene, Outcome %in% 
                           setdiff(unique(c(wbc_genes, rbc_genes, plt_genes)), ret_genes))$U_hat),
                  alternative = "greater")$statistic[['t']]
  t_rbc <- t.test(abs(filter(graph_data, Exposure == gene, Outcome %in% rbc_genes)$U_hat),
                  abs(filter(graph_data, Exposure == gene, Outcome %in% 
                           setdiff(unique(c(wbc_genes, plt_genes, ret_genes)), rbc_genes))$U_hat),
                  alternative = "greater")$statistic[['t']]
  t_wbc <- t.test(abs(filter(graph_data, Exposure == gene, Outcome %in% wbc_genes)$U_hat),
                  abs(filter(graph_data, Exposure == gene, Outcome %in% 
                           setdiff(unique(c(plt_genes, rbc_genes, ret_genes)), wbc_genes))$U_hat),
                  alternative = "greater")$statistic[['t']]
  # t_trait_non <-  t.test(abs(filter(graph_data, Exposure == gene, Outcome %in% all_blood_genes)$U_hat),
  #                        abs(filter(graph_data, Exposure == gene, !(Outcome %in% all_blood_genes))$U_hat),
  #                 alternative = "greater")$p.val
  return(data.frame(gene=gene, t_plt=t_plt, t_ret=t_ret,
              t_rbc=t_rbc, t_wbc=t_wbc)) #, p_all = p_trait_non))
}

diff_eff <- purrr::map(gene_data$gene, test_mean_effect) %>% list_rbind()
diff_eff$gene_name <- gene_data$gene_name
readr::write_csv(diff_eff, file = "/gpfs/commons/groups/knowles_lab/gwps/saved_rdata/diff_eff.csv")
diff_eff_long <- diff_eff %>% pivot_longer(-c(gene, gene_name), names_to="class", values_to = "t_val")
# diff_eff_long$p_fdr <- p.adjust(diff_eff_long$p, method='fdr')
```


```{r, fig.width=2.5, fig.height=2}
diff_eff_long %>% ggplot(aes(sample=t_val)) +
  geom_qq(distribution = qt, dparams=c("df"=250)) +
  geom_qq_line(distribution = qt, dparams=c("df"=250), line.p=c(0.1,0.9)) +
  theme_classic() + labs(x="Expected t-statistic", y="Observed t-statistic", title="Differential effects on blood trait genes") +
  theme(plot.title = element_text(size=10, hjust=1), axis.title = element_text(size=10))

```


```{r}
all_blood_genes <- union(union(plt_genes, ret_genes), union(rbc_genes, wbc_genes))
blood_ave_effect <- function(gene){
  plt_effect <- mean(abs(filter(graph_data, Exposure == gene, Outcome %in% plt_genes)$U_hat))
  ret_effect <- mean(abs(filter(graph_data, Exposure == gene, Outcome %in% ret_genes)$U_hat))
  wbc_effect <- mean(abs(filter(graph_data, Exposure == gene, Outcome %in% wbc_genes)$U_hat))
  rbc_effect <- mean(abs(filter(graph_data, Exposure == gene, Outcome %in% rbc_genes)$U_hat))
  non_trait_effect <- mean(abs(filter(graph_data, Exposure == gene, Outcome %in% all_blood_genes)$U_hat))
  return(data.frame("gene" = gene, "PLT" = plt_effect, "RET" = ret_effect,
                    "WBC" = wbc_effect, "RBC" = rbc_effect, "None" = non_trait_effect))
}

gene_blood_ave_eff <- purrr::map(gene_data$gene, blood_ave_effect) %>% list_rbind()
gene_blood_ave_eff$gene_name <- gene_data$gene_name
```


```{r, fig.width=3.5, fig.height=2}
gene_blood_ave_eff %>% select(-starts_with('gene')) %>% cor %>% as_tibble(rownames="index") %>% pivot_longer(-index, values_to="Cor") %>%
  mutate(CorP = 100*Cor) %>%
  ggplot(aes(x=index,y=name)) + geom_raster(aes(fill=CorP)) + scale_fill_gradient2(limits=c(0,100)) +
  geom_text(aes(index, name, label=round(CorP, 2)), colour = "white", check_overlap = TRUE, size=3.5) + theme_minimal() +
  labs(x=NULL, y=NULL, fill="Cor %") + theme(legend.key.size = unit(0.4, 'cm'), legend.title = element_text(size=10))
```


# ```{r}
# cutoff <- 10
# options(digits=3)
# gene_blood_ave_eff %>% mutate(rnk_plt = min_rank(desc(mean_plt_eff)),
#                               rnk_ret = min_rank(desc(mean_ret_eff)),
#                               rnk_rbc = min_rank(desc(mean_rbc_eff)),
#                               rnk_wbc = min_rank(desc(mean_wbc_eff))) %>%
#   filter((rnk_wbc <= cutoff)|(rnk_rbc <= cutoff)|(rnk_ret <= cutoff)|(rnk_plt <= cutoff)) %>%
#   arrange(rnk_plt+rnk_ret+rnk_ret+rnk_wbc) %>%
#   select(gene_name, ends_with('eff'))
# ```
# 
# 
# 
# ```{r}
# cut = 0.1
# 
# plt_high_genes <- unique(filter(graph_data, Exposure != Outcome, Outcome %in% plt_genes, abs(U_hat) > 0.1)$Exposure)
# ret_high_genes <- unique(filter(graph_data, Exposure != Outcome, Outcome %in% ret_genes, abs(U_hat) > 0.1)$Exposure)
# rbc_high_genes <- unique(filter(graph_data, Exposure != Outcome, Outcome %in% rbc_genes, abs(U_hat) > 0.1)$Exposure)
# wbc_high_genes <- unique(filter(graph_data, Exposure != Outcome, Outcome %in% wbc_genes, abs(U_hat) > 0.1)$Exposure)
# 
# blood_high_genes <- union(union(plt_high_genes, ret_high_genes), union(rbc_high_genes, wbc_high_genes))
# ```


```{r}
trait_genes <- list('plt'=plt_genes, 'ret'=ret_genes, 'wbc'=wbc_genes, 'rbc'=rbc_genes)

calc_trait_distances <- function(trait1, trait2){
  subset_wi <- filter(graph_data, Exposure %in% trait_genes[[trait1]], Outcome %in% trait_genes[[trait1]])
  if(trait2 != 'all'){
    subset_bt <- filter(graph_data, (Exposure %in% trait_genes[[trait1]] & Outcome %in% trait_genes[[trait2]])|(Exposure %in% trait_genes[[trait2]] & Outcome %in% trait_genes[[trait1]]))
  }
  else{
    subset_bt <- filter(graph_data, !(Exposure %in% trait_genes[[trait1]]), !(Outcome %in% trait_genes[[trait1]]))
  }
  t_test_res <- t.test(subset_wi$path_length, subset_bt$path_length)
  return(data.frame("trait1" = trait1, "trait2" = trait2,
                    "path_length" = mean(subset_bt$path_length, na.rm=T),
                    "sd" = sd(subset_bt$path_length, na.rm=T),
                    "se" =sd(subset_bt$path_length, na.rm=T)/sqrt(length(subset_bt$path_length)),
                    p = t_test_res$p.val))
}

grid <- expand.grid(trait1=names(trait_genes), trait2=c(names(trait_genes), 'all')) 

bt_wi_distance <- purrr::map2(grid$trait1, grid$trait2, calc_trait_distances) %>% purrr::list_rbind()
bt_wi_distance$p_adj <- p.adjust(bt_wi_distance$p)
```


```{r}
options(scipen=2)
options(digits=4)
bt_wi_distance %>% filter(trait2 != 'all') %>% arrange(trait1) %>% print(n=20)
```

